---
title: "GIS Final Exam"
author: "Stephen Lewandowski"
date: "December 10, 2018"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(maptools)
library(rgdal)
library(spgwr)
library(spdep)

knitr::opts_chunk$set(echo = TRUE)
```


Read in the "US_CTY_PreventiveHealthExp.shp" file to create the "health_exp" object.

```{r import_shp}

health_exp <- readOGR(dsn = "./shapefile", layer = "US_CTY_PreventativeHealthExp")

names(health_exp)
plot(health_exp)

```



## Geographically Weighted Regression (GWR)

Effect of %BA on expenditure

```{r gwr}
bwG <- gwr.sel(PHlth_Exp ~ PCT_BA + PCT_RUR + PCT_POV + PCT_WH, data = health_exp, gweight = gwr.Gauss, verbose = TRUE)

gwrG <- gwr(PHlth_Exp ~ PCT_BA + PCT_RUR + PCT_POV + PCT_WH, data = health_exp, bandwidth = bwG, gweight = gwr.Gauss)


```


Plot of the spatially varying effect of % with BA on expenditure across the country.
```{r plot_effect_BA}
spplot(gwrG$SDF, "PCT_BA")
```
The effect of percent with a BA degree on expenditure is weaker in the areas closer to zero and stronger in the areas further from zero. 



Plot of the local R2 of GWR model.
```{r plot_localR2}
#plot r-square
spplot(gwrG$SDF, "localR2")
```
These results indicate that the full specified model fits best (highest R-square) in certain regions of the country such as the southwest/California, southern Florida, western Washington, and Maine. The model fits poorly in parts of the great plains/ central United States, along the Mississippi River, and in Oregon, amongst other places. 


Create shapefile with GWR results.
```{r write_gwr_shapefile}
#create .shp
writeSpatialShape(gwrG$SDF, "GWR_Results")
```


## Moran's I scatterplots

Import data.

```{r import_geoda}
geoda <- read_csv("../final_geoda.csv")
```

###Expenditures cluster analysis, Moran's I
```{r cluster_morans_i}

geoda %>% 
ggplot(aes(x = MORAN_STD, y = MORAN_LAG)) + 
  geom_point() + 
  geom_smooth(method = lm) +
  labs(
    title = "Expenditures Moran's I: 0.765772",
    x = "Preventative Health Expenditures",
    y = "Lagged Preventative Health Expenditures",
    caption = "p = 0.001"
    ) +
  theme_bw()
```

###OLS Residuals Moran's I
```{r ols_resid_morans_i}

geoda %>% 
ggplot(aes(x = ERR_STD, y = ERR_LAG)) + 
  geom_point() + 
  geom_smooth(method = lm) +
  labs(
    title = "Spatial Error Residuals Moran's I: -0.166303",
    x = "ERR Residuals",
    y = "Lagged ERR Residuals",
    caption = "p = 0.001"
    ) +
  theme_bw()
```


###Percent BA Moran's I
```{r err_resid_morans_i}

geoda %>% 
ggplot(aes(x = MOR_STD_BA, y = MOR_LAG_BA)) + 
  geom_point() + 
  geom_smooth(method = lm) +
  labs(
    title = "Percent BA Moran's I: 0.39285",
    x = "Percent BA",
    y = "Lagged Percent BA",
    caption = "p = 0.001"
    ) +
  theme_bw()
```

## Weight Matrix histogram

```{r histogram}
geoda %>% 
ggplot(aes(NUM_NBRS)) + 
  geom_histogram() + 
    labs(
    title = "U.S. county weight matrix histogram",
    x = "Number of Neighbors",
    y = "count"
    ) +
  scale_x_continuous(breaks = seq(0,11,1)) +
  theme_bw()
```

